---
output: 
  html_document: 
    keep_md: yes
---
```{r echo=FALSE}
options(width=120)
knitr::opts_knit$set(verbose = TRUE)
knitr::opts_chunk$set(cache=TRUE)
```
Statistic tests
======================================


### statistic tests about sequences coming from ribosomal RNA

Regarding the PS and RanN6 set, we use a paired t.test as the the results come from 3 independents experiments.

```{r}
rDNA <- read.table('rDNA.csv', sep=",", head=T)
rDNA
```

```{r}
t.test(rDNA$HeLa_PS, rDNA$HeLa_RanN6, paired = T)
```

```{r}
t.test(rDNA$THP1_PS, rDNA$THP1_RanN6, paired = T)
```

```{r}
rDNA_40N6 <- read.table('rDNA_40N6.csv', sep=",", head=T)
rDNA_40N6
```

Regarding the 40N6 set, we can not use the paired test as only one experiment has been performed. Thus, we use the 3 replicats of 1 experiment.

```{r}
t.test(rDNA_40N6$HeLa_RanN6, rDNA_40N6$HeLa_40N6)
```

```{r}
t.test(rDNA_40N6$HeLa_PS, rDNA_40N6$HeLa_40N6)
```

```{r}
t.test(rDNA_40N6$THP1_RanN6, rDNA_40N6$THP1_40N6)
```

```{r}
t.test(rDNA_40N6$THP1_PS, rDNA_40N6$THP1_40N6)
```

### statistic tests about sequences coming from artefacts

Regarding the PS and RanN6 set, we use a paired t.test as the the results come from 3 independents experiments.

```{r}
artefact <- read.table('artefacts.csv', sep=",", head=T)
artefact
```

```{r}
t.test(artefact$HeLa_PS, artefact$HeLa_RanN6, paired = T)
```

```{r}
t.test(artefact$THP1_PS, artefact$THP1_RanN6, paired = T)
```

Regarding the 40N6 set, we can not use the paired test as only one experiment has been performed. Thus, we use the 3 replicats of 1 experiment.

```{r}
artefact_40N6 <- read.table('artefact_40N6.csv', sep=",", head=T)
artefact_40N6
```

```{r}
t.test(artefact_40N6$HeLa_RanN6, artefact_40N6$HeLa_40N6)
```

```{r}
t.test(artefact_40N6$HeLa_PS, artefact_40N6$HeLa_40N6)
```

```{r}
t.test(artefact_40N6$THP1_RanN6, artefact_40N6$THP1_40N6)
```

```{r}
t.test(artefact_40N6$THP1_PS, artefact_40N6$THP1_40N6)
```


### statistic tests about the numbers of genes detected

Regarding the PS and RanN6 set, we use a paired t.test as the the results come from 3 independents experiments.

```{r}
genes <- read.table('genes.csv', sep=",", head=T)
genes
```

```{r}
t.test(genes$HeLa_PS, genes$HeLa_RanN6, paired = T)
```

```{r}
t.test(genes$THP1_PS, genes$THP1_RanN6, paired = T)
```

Regarding the 40N6 set, we can not use the paired test as only one experiment has been performed. Thus, we use the 3 replicats of 1 experiment.

```{r}
genes_40N6 <- read.table('genes_40N6.csv', sep=",", head=T)
genes_40N6
```

```{r}
t.test(genes_40N6$HeLa_RanN6, genes_40N6$HeLa_40N6)
```

```{r}
t.test(genes_40N6$HeLa_PS, genes_40N6$HeLa_40N6)
```

```{r}
t.test(genes_40N6$THP1_RanN6, genes_40N6$THP1_40N6)
```

```{r}
t.test(genes_40N6$THP1_PS, genes_40N6$THP1_40N6)
```
